### 2022-06-04 src/ms.tex
finish writing the occultation section of chapter 2

% Math 
\usepackage{amsmath,amssymb,bm}
\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{siunitx} % for proper units

% Other packages
\usepackage{url}
\usepackage{epstopdf}
\usepackage[font=scriptsize]{caption} % smaller font for fig captions
\usepackage{placeins}
%\usepackage{import}
% of references

@@

Later works such as
\citet{1992Icar...97..211B,1993Icar..102..134Y,1999AJ....117.1063Y} went a step
further by using not only phase curves but also light curves of mutual
\emph{occultations} of Pluto by its moon Charon to reconstruct albedo maps with
greater success. The major advantage of occultations relative to just phase
curves is that more information about the surface is encoded into the light
curve because of the sharp limb of the occultor sweeping over the disc of an
occulted body and blocking the reflected light. More importantly for this
thesis, \citet{1994Icar..107..195S} was the first to observe the occultations
of Jupiter's moon Io by another of Jupiter's moons -- Europa, and also
occultations of Io by Jupiter itself. These observations were conducted using
near-infrared telescopes such as NASA's Infrared Telescope Facility (IRTF) to
observe emitted light from Io's surface which is covered with many time-varying
and bright volcanic features. The observing campaign of Io has yielded insights
into the the nature of its volcanic activity and it continues to this day. I
will return to the subject of Io in great detail in
Chapter~\ref{ch:mapping_io}.

@@

\subsection{Magnification of an extended source}
For a relatively small subset of microlensing events the source star passes
very close to the caustic and the variation of magnification over the extent of
the source star disc is non-negligible. For those events the point source
approximation breaks down and we have to take into account the finite angular
radius $\theta_\star$. This will generally happen when $u_0 \lesssim
    \rho_\star/2$ \citep{1997ApJ...477..580G} where
$\rho_\star=\theta_\star/\theta_E$. The effect of finite source effects on the
magnification curve thus matters only near the peak of an event and it results
in a rounder top of the curves shown in Figure~\ref{fig:paczynski_curve}.
Unfortunately there are no analytic solutions for the magnification of an
extended source lensed by a single point lens with a general surface brightness
profile. \citet{1994ApJ...421L..71G} derived an approximate solution valid for
uniform brightness sources with $\rho_\star\lesssim 0.1$ and
\citet{1994ApJ...430..505W} derived a general solution (also for a uniform
brightness source) involving solutions to elliptic integrals of the first,
second and third kind. A general solution involving Fast Fourier Transforms
which is applicable to any source profile was recently proposed by
\citet{2022arXiv220306637S}.

\begin{figure}[t]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/single_lens_images.pdf}
        \caption{Images of an extended limb-darkened source star lensed by a single
            point lens for varying position of the source star. The top panels show the magnification map with a logarithmic scale
            colormap  consisting of a single point caustic. The semi-transparent circle
            is the source star disc with radius $\rho_\star=0.15$. The bottom row shows
            the two images merging into the Einstein ring as the source moves over the caustic.}
        \label{fig:single_lens_images}
    \end{centering}
\end{figure}

@@

\subsection{Binary lens}
In this section we will focus specifically on the binary lens. We choose a
coordinate system whose origin is at the midpoint of the line connecting the
two lenses and both lenses are on the real axis. If the first lens is at
distance $a$ ($z_1=a$) then the second lens is at $-a$ and the lens equation
takes the form
\begin{equation}
    w=z+\frac{\epsilon_{1}}{\bar{z} - a}+\frac{1 - \epsilon_{1}}{\bar{z} + a}
\end{equation}
The above equation can be rewritten as a 5th order complex polynomial
(see Appendix~\ref{app:complex_poly}). Either 3 or 5 of the complex roots of that
polynomial are also solutions to the lens equation
(5 inside the caustics and 3 outside).
It is common to use the mass ratio $q\equiv \epsilon_2/\epsilon_1$ with
$\epsilon_2 \leq \epsilon_1$ instead  of $\epsilon_1$ as a parameter and the
separation between the lenses $s\equiv 2a$ instead of half the separation $a$.
As a reminder all angular quantities are expressed in the units of the angular Einstein
radius of the total mass of the system.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/binary_lens_topology.pdf}
    \caption{The three different topologies of the binary lens. The left panels show the
        critical curves in the lens plane, the right panels show the caustic curves in the
        source plane. The figure shows all three topologies and the transitions between
        them for an equal mass binary lens with $q=1$. In this case the transition between
        the close and intermediate topology occurs at $s=1/\sqrt{2}$ and the one between
        the intermediate and wide topology at $s=2.0$. Figure adapted from
        \citet{dominik1999}.
    }
    \label{fig:binary_lens_topology}
\end{figure}

@@

          \citet{2003ApJ...586..527G} focused specifically on the
          possibility of inferring

@@

\section{Occultation and phase curve mapping}
\label{sec:occultations}
\subsection{The Starry framework}
\subsection{Reflected vs. emitted light}

% In this and the following sections make sure to use microlensing and occultation
% mapping as examples
\section{Bayesian statistics}
\label{sec:bayes}
\subsection{Probability theory}
\subsection{The meaning of probability}
\subsection{Probability as frequency of events in repeated trials}
\subsection{Probability as degrees of belief}
\subsection{The likelihood function}
% mention the likelihood principle
\subsection{Priors}
\subsection{On machine learning and the difference between explanation and prediction}
% mention that even if we have a model focused on inference, it might be best if we judge 
% its quality through its predictive performance

\section{Inference}
\label{sec:inference}
% check that paper on physics and ML by that MIT guy, there's a good stuff there. 
% check also Ian Murray's lecture notes  on MLPR
\subsection{The curse of dimensionality}
\subsection{There's no free lunch}
% exploration vs. explotation 
\subsection{Optimization vs. sampling}
\subsection{Sampling and optimization methods}
\subsubsection{Maximum likelihood}
\subsubsection{The Laplace approximation}
\subsubsection{Variational inference}
\subsubsection{Importance sampling}
\subsubsection{Rejection sampling}
\subsubsection{Markov Chain Monte Carlo}
\subsubsection{Hamiltonian Monte Carlo}
\subsubsection{Other notable methods}

>>>>

% Math 
\usepackage{amsmath,amssymb,bm}
\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{siunitx} % for proper units 
\usepackage{esint} % various fancy integral symbols
\usepackage[bbgreekl]{mathbbol}

% Other packages
\usepackage{url}
\usepackage{epstopdf}
\usepackage[font=scriptsize]{caption} % smaller font for fig captions
\usepackage[section]{placeins}
%\usepackage{import}
% of references

@@

Later works such as
\citet{1986AJ.....92.1201D,1992Icar...97..211B,1993Icar..102..134Y,1999AJ....117.1063Y}
went a step further by using not only phase curves but also light curves of
mutual \emph{occultations} of Pluto by its moon Charon to reconstruct albedo
maps with greater success. The major advantage of occultations relative to just
phase curves is that more information about the surface is encoded into the
light curve because of the sharp limb of the occultor sweeping over the disc of
an occulted body and blocking the reflected light. More importantly for this
thesis, \citet{1994Icar..107..195S} was the first to observe the occultations
of Jupiter's moon Io by another of Jupiter's moons -- Europa, and also
occultations of Io by Jupiter itself. These observations were conducted using
near-infrared telescopes such as NASA's Infrared Telescope Facility (IRTF) to
observe emitted light from Io's surface which is covered with many time-varying
and bright volcanic features. The observing campaign of Io has yielded insights
into the the nature of its volcanic activity and it continues to this day. I
will return to the subject of Io in great detail in
Chapter~\ref{ch:mapping_io}.

@@

\subsection{Magnification of an extended source}
\begin{figure}[t]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/single_lens_images.pdf}
        \caption{Images of an extended limb-darkened source star lensed by a single
            point lens for varying position of the source star. The top panels show the magnification map with a logarithmic scale
            colormap  consisting of a single point caustic. The semi-transparent circle
            is the source star disc with radius $\rho_\star=0.15$. The bottom row shows
            the two images merging into the Einstein ring as the source moves over the caustic.}
        \label{fig:single_lens_images}
    \end{centering}
\end{figure}
For a relatively small subset of microlensing events the source star passes
very close to the caustic and the variation of magnification over the extent of
the source star disc is non-negligible. For those events the point source
approximation breaks down and we have to take into account the finite angular
radius $\theta_\star$. This will generally happen when $u_0 \lesssim
    \rho_\star/2$ \citep{1997ApJ...477..580G} where
$\rho_\star=\theta_\star/\theta_E$. The effect of finite source effects on the
magnification curve thus matters only near the peak of an event and it results
in a rounder top of the curves shown in Figure~\ref{fig:paczynski_curve}.
Unfortunately there are no analytic solutions for the magnification of an
extended source lensed by a single point lens with a general surface brightness
profile. \citet{1994ApJ...421L..71G} derived an approximate solution valid for
uniform brightness sources with $\rho_\star\lesssim 0.1$ and
\citet{1994ApJ...430..505W} derived a general solution (also for a uniform
brightness source) involving solutions to elliptic integrals of the first,
second and third kind. A general solution involving Fast Fourier Transforms
which is applicable to any source profile was recently proposed by
\citet{2022arXiv220306637S}.

@@

\subsection{Binary lens}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/binary_lens_topology.pdf}
    \caption{The three different topologies of the binary lens. The left panels show the
        critical curves in the lens plane, the right panels show the caustic curves in the
        source plane. The figure shows all three topologies and the transitions between
        them for an equal mass binary lens with $q=1$. In this case the transition between
        the close and intermediate topology occurs at $s=1/\sqrt{2}$ and the one between
        the intermediate and wide topology at $s=2.0$. Figure adapted from
        \citet{dominik1999}.
    }
    \label{fig:binary_lens_topology}
\end{figure}
In this section we will focus specifically on the binary lens. We choose a
coordinate system whose origin is at the midpoint of the line connecting the
two lenses and both lenses are on the real axis. If the first lens is at
distance $a$ ($z_1=a$) then the second lens is at $-a$ and the lens equation
takes the form
\begin{equation}
    w=z+\frac{\epsilon_{1}}{\bar{z} - a}+\frac{1 - \epsilon_{1}}{\bar{z} + a}
\end{equation}
The above equation can be rewritten as a 5th order complex polynomial
(see Appendix~\ref{app:complex_poly}). Either 3 or 5 of the complex roots of that
polynomial are also solutions to the lens equation
(5 inside the caustics and 3 outside).
It is common to use the mass ratio $q\equiv \epsilon_2/\epsilon_1$ with
$\epsilon_2 \leq \epsilon_1$ instead  of $\epsilon_1$ as a parameter and the
separation between the lenses $s\equiv 2a$ instead of half the separation $a$.
As a reminder all angular quantities are expressed in the units of the angular Einstein
radius of the total mass of the system.

@@

          \citet{2003ApJ...586..527G} focused specifically on the possibility of inferring

@@

\section{Occultation and phase curve mapping}
\label{sec:occultations}
In this section I will briefly review the theory behind modeling phase curves and
occultation light curves of spherical bodies for the case of emitted light
(isotropic blackbody emission from the body itself) and reflected light
(starlight scattered from the surface or atmosphere of the body). The following is
mostly a short summary of the papers introducing the \textsf{starry} framework
\citep{2019AJ....157...64L,2021arXiv210306275L}.
\subsection{The Starry framework}
\begin{figure}[t]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/spherical_harmonics.pdf}
        \caption{Real spherical harmonics up to degree $l=5$ computed from
            Equation~\ref{eq:spherical_harmonics}. Figure adapted from Figure 1 in
            \citet{2019AJ....157...64L}.}
        \label{fig:spherical_harmonics}
    \end{centering}
\end{figure}
\subsubsection{Emitted light}
Consider a spherical body of unit radius emitting light isotropically (a
Lambertian emitter) at each point $(\theta,\phi)$ where $\theta$ is the
inclination angle and $\phi$ is the azimuthal angle. To compute a phase curve
or an occultation light curve we need to be able to integrate the flux over the
entire projected disc of the body given the distribution of specific intensity
$I(\theta, \phi)$, the 3D orientation of the body in space, and the location of
a possible occultor relative to the body. These two-dimensional integrals can
always be computed numerically (see for instance the approaches presented in
\citet{2018AJ....156..146F,2018MNRAS.477.2613L}) but
\citet{2019AJ....157...64L} showed that it is possible to compute them
analytically if the specific intensity is first expanded in a basis of
spherical harmonics and then mapped to a different basis more suitable for
evaluating the integrals.

Following \citet{2019AJ....157...64L}, we start by setting up a cartesian
coordinate system on the units sphere, such that
\begin{align}
     & x=\sin \theta \cos \phi \\
     & y=\sin \theta \sin \phi \\
     & z=\cos \theta
\end{align}
and the observer is located on the $z$-axis at $\infty$ such that the projected disc
of the body is centered at the origin of the $xy$ plane with the unit vector
$\hat{\mathbf{x}}$ pointing to the right and $\hat{\mathbf{y}}$ pointing up.
We introduce spherical harmonics $Y_{l m}(\theta, \phi)$ of degree $l\geq 0$ and order
$m\in [-l, l]$ defined  as
\begin{equation}
    Y_{l m}(\theta, \phi)= \begin{cases}\bar{P}_{l m}(\cos \theta) \cos (m \phi) & m \geqslant 0 \\ \bar{P}_{l|m|}(\cos \theta) \sin (|m| \phi) & m<0\end{cases}
    \label{eq:spherical_harmonics}
\end{equation}
When the spherical harmonics defined above are rewritten in terms of $x$, $y$, and $z$
they become polynomials in these variables (see Appendix A in \citet{2019AJ....157...64L}).
We can expand the specific intensity distribution $I(x,y)$ in a spherical harmonic basis
as
\begin{equation}
    I(x, y)=\tilde{\mathbf{y}}^{\top}(x, y) \mathbf{y}
    \label{eq:sh_expansion}
\end{equation}
where $\tilde{\mathbf{y}}$ is the basis vector of spherical harmonics  arranged in
increasing order:
\begin{align}
     & \tilde{\mathbf{y}}=\left(\begin{array}{lllll}
                                    Y_{0,0} & Y_{1,-1} & Y_{1,0} & Y_{1,1} & Y_{2,-2}
                                \end{array}\right. \\
     & \left.\begin{array}{lllll}
                 Y_{2,-1} & Y_{2,0} & Y_{2,1} & Y_{2,2} & \cdots
             \end{array}\right)^{\top} \quad,
\end{align}
and $\mathbf{y}$ is a vector of scalar \emph{spherical harmonic coefficients}.
$\mathbf{y}$ is a central quantity which defines the \emph{map} (the specific
intensity at every point) of the body.
\citet{2019AJ....157...64L} shows that we can represent $y$ in a polynomial basis (see
\citet{2019AJ....157...64L} for the general expression)
\begin{equation}
    \tilde{\boldsymbol{p}}=\left(\begin{array}{llllllllll}
        1 & x & z & y & x^{2} & x z & x y & y z & y^{2} & \cdots
    \end{array}\right)^{\top}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{p}                    \\
            & =\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_1 \mathbf{y}\quad,
    \label{eq:intensity_poly_basis}
\end{align}
and also in the Green's basis
\begin{equation}
    \tilde{\boldsymbol{g}}=\left(\begin{array}{llllllllll}
        1 & 2 x & z & y & 3 x^{2} & -3 x z & 2 x y & 3 y z & y^{2} & \cdots
    \end{array}\right)^{\top}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{g}              \\
            & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{A}_2 \mathbf{p} \\
            & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{A} \mathbf{y}
\end{align}
where $\mathbf{A}_1$ is the change-of-basis matrix from $\mathbf{y}$ to $\mathbf{p}$,
$\mathbf{A}_2$ is the change-of-basis matrix from $\mathbf{p}$ to $\mathbf{g}$, and
$\mathbf{A}\equiv \mathbf{A}_1 \mathbf{A}_2$.

To compute rotational light curves we also need to be able to specify the
orientation of a surface map with coefficients $\mathbf y$. We can rotate the
map using the Wigner rotation matrix $\mathbf{R}(\mathrm{I}, \Lambda, \Theta)$
that rotates $y$ given the body's inclination $\mathrm{I}$, obliquity $\Lambda$
and rotational phase $\Theta$. The rotated map is then $\mathbf{R}(\mathrm{I},
    \Lambda, \Theta) \mathbf{y}$ and Equation~\ref{eq:intensity_poly_basis} can be
rewritten more generally as
\begin{equation}
    I(x, y)=\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_1 \mathbf{R}\mathbf{y}
\end{equation}
The total flux measured by an observer is is given by an integral of the specific
intensity over a region S of the projected disc of the body:
\begin{align}
    F & =\oiint I(x, y) \ud S                                                            \\
      & =\oiint \tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_{1} \mathbf{R} \mathbf{y} d S \\
      & =\mathbf{r}^{\top} \mathbf{A}_{1} \mathbf{R} \mathbf{y}
\end{align}
where $\mathbf{r}$ is a column vector whose $n$-th element is \citep{2019AJ....157...64L}
\begin{equation}
    r_{n} \equiv \oiint \tilde{p}_{n}(x, y) \ud S
\end{equation}
When the entire disc is visible (no occultor)
\begin{equation}
    r_{n}=\int_{-1}^{1} \int_{-\sqrt{1-x^{2}}}^{\sqrt{1+x^{2}}} \tilde{p}_{n}(x, y) \mathrm{d}y \mathrm{d} x
\end{equation}
and the solution to this integral can be expressed in terms of Gamma functions
\citep[Equation 20 in ][]{2019AJ....157...64L}.
$\mathbf{r}$ and $\mathbf{A}_1$ are independent of the map coefficients $\mathbf{y}$
so they can be pre-computed.

Computing the occultation light curves is more complicated because the the body
is partially occulted by an occultor of radius $r$, centered at $(x_o, y_o)$
and the exposed portion of the disc is a function of $(r, x_o, y_o)$. The
general expression for the flux is
\begin{align}
    F & =\oiint I(x, y) \mathrm{d} S                         \\
      & =\mathbf{s}^\top  \mathbf{A} \mathbf{R} \mathbf{y} .
\end{align}
where the vector
\begin{equation}
    \mathbf{s}^\top \equiv \oiint \tilde{\mathbf{g}}^{\top}(x, y) \mathrm{d} S
\end{equation}
is defined to be the solution to the integral.
\citet{2019AJ....157...64L} solves this integral in two steps.  First, we rotate
the coordinate system about the  $z$-axis  so that the occultor lies at along the
$+y$ axis and its center is at
a distance $b=\sqrt{x_o^{2}+y_o^{2}}$ from the origin. This substantially simplifies
the limits of integration.
The second step is to use Green's theorem to express the surface integral of
$\tilde{\mathbf{g}}_n$ as a line integral of a vector function $\mathbf{G}_n$
along the boundary of the visible portion of the occulted disc.
The $n$-th element of the solution vector $\mathbf{s}^\top$ is
\begin{equation}
    s_{n}=\oiint \tilde{g}_{n}(x, y) \mathrm{d} S=\oint \mathbf{G}_{n}(x, y) \cdot \ud \mathbf{r}
\end{equation}
where
$\mathbf{G}_{n}(x, y)=G_{n x}(x, y) \hat{\mathbf{x}}+G_{n y}(x, y) \hat{\mathbf{y}}$
is constructed such that
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n}=\tilde{g}_{n}(x, y)
    \label{eq:G_n}
\end{equation}
were $\mathbf{D} \wedge \mathbf{G}_{n}$ is the exterior derivative of $\mathbf{G}_n$
given by
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n} \equiv \frac{\ud G_{n y}}{\ud x}-\frac{\ud G_{n_{x}}}{\ud y}
\end{equation}
in Cartesian coordinates.
\citet{2019AJ....157...64L} provides one possible expression for $\mathbf{G}_n$
which satisfies Equation~\ref{eq:G_n} and shows that the final solution can be written as
\begin{equation}
    s_{n}=\mathcal{Q}\left(\mathbf{G}_{n}\right)-\mathcal{P}\left(\mathbf{G}_{n}\right)
\end{equation}
where $\mathcal{P}(\mathbf{G}_n)$ is the  line integral along the arc of the occultor
of radius $r$ and $\mathcal{Q}(\mathbf{G}_n)$ is the line integral along the arc of the
occulted body of unit radius. Solutions to these integrals are long but they consist of
analytic functions such as sines, cosines and complete elliptic integrals and importantly
they only need to be evaluated once for an arbitrary map with fixed geometry of an
occultation \citep{2019AJ....157...64L}.

The final expression for the total flux for a particular geometrical
arrangement of the occulted body and the occultor can then be written
succinctly as
\begin{equation}
    f = \mathbf{s}^T\mathbf{A}\mathbf{R}'\mathbf{R}\mathbf{y}
    \label{eq:starry}
\end{equation}
Where $\mathbf{R}'$ is a rotation matrix which rotates the map such that the occultor
is placed symmetrically on the $+y$ axis at distance $b$.
Notice that Equation~\ref{eq:starry} defines a linear operation acting on a vector
of spherical harmonic coefficients $\mathbf{y}$.
It is also trivial to generalize this expression to a case of \emph{spectral map}. In
that case we define a matrix $Y$ whose columns are the spherical harmonic coefficients
at different wavelengths and the total flux (spectrum) is then
\begin{equation}
    \mathbf{f} = \mathbf{s}^T\mathbf{A}\mathbf{R}'\mathbf{R}\mathbf{Y}
    \label{eq:starry_mw}
\end{equation}
where $\mathbf{f}$ is a vector of fluxes, one per wavelength bin.

The formalism above can also be extended to account for limb-darkening which is
important when modeling transits of planets across stars or the light curves of
eclipsing binaries. \citet{2020AJ....159..123A} show how a limb-darkening
profile that is an order $l$ polynomial function of
$\mu=z=\sqrt{1-x^{2}-y^{2}}$ can be expressed in terms of the $m=0$ spherical
harmonics up to order $l$. In the case of quadratic limb-darkening
$(l_\mathrm{max}=2)$, the limb darkening profile is
\begin{equation}
    \frac{I(\mu)}{I(1)}=1-u_{1}(1-\mu)-u_{2}(1-\mu)^{2}
\end{equation}
which can be expressed as a sum of spherical harmonics
\citep[Equation 38 in][]{2019AJ....157...64L}. Limb-darkening needs to be
treated separately from the map coefficients $\mathbf{y}$ because it does not rotate along with the map
when rotations $\mathbf{R}$ and $\mathbf{R}'$ are applied. The limb-darkening
coefficients are applied to the map $\mathbf{y}$ as a multiplicative filter
following any rotations which results in another set of spherical harmonic
coefficients because products of spherical harmonics are also spherical
harmonics. Applying limb-darkening raises the degree of the map by the degree
of the limb-darkening.

\subsubsection{Reflected light}
The results derived in the previous section are valid only for light emitted
from the occulted body. Equation~\ref{eq:starry} is a good model for stellar
light curves and secondary eclipse light curves and phase curves of planets
observed in the mid to far infrared wavelengths where the contribution of
scattered starlight to total flux is negligible. Modeling reflected light phase
curves and occultations is far more complicated because one has to take into
account the the nonuniform illumination of the body and the possible presence
of a terminator line (boundary between day and night).
\citet{2021arXiv210306275L} extends the formalism described in the previous
section to the case of reflected light phase curves and occultations for a body
whose spatial \emph{albedo} (specifically the \emph{spherical albedo} -- the
fraction of power incident on a body at a given wavelength that is scattered
into space in all directions) distribution is expanded in a basis of spherical
harmonics. Thus, analogous to Equation~\ref{eq:sh_expansion} we have
\begin{equation}
    A(x, y)=\tilde{\mathbf{y}}^{\top}(x, y) \mathbf{y}
\end{equation}
where $A$ is the albedo. We work under the assumption of isotropic (Lambertian)
scattering of light at every point on the body which means that the illumination
profile $\mathcal{I}$ of the surface is given by Lambert's law:
\begin{equation}
    \mathcal{I}\left(\vartheta_{\mathrm{i}}\right)=\mathcal{I}_{0} \max \left(0, \cos \vartheta_{\mathrm{i}}\right)
\end{equation}
where $\vartheta_i$ is the angle between the incident radiation at the surface normal
and $\mathcal{I}_0$ is the maximum illumination. In case of, for example, a planet illuminated
by its host star \citep[Appendix A.2 in][]{2021arXiv210306275L}
\begin{equation}
    \mathcal{I}_{0}=\frac{f_{s}}{\pi r_{\mathrm{s}}^{2}}
\end{equation}
where $r_s$ is the distance between the planet and the star in units of planet's radius
and $f_s$ is the stellar flux measured at the observer in some arbitrary units.
We assume that $f_s=1$ so that all fluxes are defined as the fraction of the flux
of the illumination source at the observer.
It is the presence of the day/night terminator that complicates the calculation
of total flux the most. The limits of integration end up depending on a
solution to a quartic equation specifying the points of intersection between
the occultor and the day/night terminator line and the solution to those
integrals are a function of \emph{incomplete} elliptic integrals as opposed to
complete elliptic integrals as was the case for occultations in emitted light.
The full derivation of the total flux is provided by \citet{2021arXiv210306275L} and
I will only briefly summarize  the result here for completeness.

Under the Lambertian scattering assumption the observed intensity at any point
of the surface of the body is is proportional to the cosine of the angle
$\vartheta_i$ between the incident light rays and the surface normal. All
points for which $\vartheta_i\geq \pi/2$ have intensity of zero (they're
unilluminated). Let's assume that the point source is placed at sky coordinates
$\left(x_{\mathrm{s}}, y_{\mathrm{s}}, z_{\mathrm{s}}\right)$ in units of the
radius of the illuminated body. The day/night terminator on the body is a
half-ellipse with a semi-major axis equal to unity and a (signed) semi-minor
axis equal to
\begin{equation}
    b=-\frac{z_{\mathrm{s}}}{r_{\mathrm{s}}}
\end{equation}
where $r_{\mathrm{s}}=\sqrt{x_{\mathrm{s}}^{2}+y_{\mathrm{s}}^{2}+z_{\mathrm{s}}^{2}}$
is the distance to the source. The angle by which the semi-major axis of this ellipse is
rotated away from the $+x$-axis is  is
\begin{equation}
    \theta=-\arctan 2\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right)
\end{equation}
Under the assumption that $rs \gg 1$, the illumination $\mathcal{I}$ at point
$(x,y)$ on the projected disc is \citep{2021arXiv210306275L}
\begin{equation}
    \mathcal{I}\left(b, \theta, r_{\mathrm{s}} ; x, y\right)=\max \left(0, I\left(b, \theta, r_{\mathrm{s}} ; x, y\right)\right)
    \label{eq:illumination}
\end{equation}
where
\begin{align}
    I\left(b, \theta, r_{\mathrm{s}} ; x, y\right) & =\frac{1}{\pi r_{\mathrm{s}}^{2}} \cos \vartheta_{\mathrm{i}}                                                      \\
                                                   & =\frac{1}{\pi r_{\mathrm{s}}^{2}}\left(-b_{\mathrm{c}} \sin \theta x+b_{\mathrm{c}} \cos \theta y-b z(x, y)\right)
\end{align}
with  $b_{\mathrm{c}} \equiv \sqrt{1-b^{2}}$  and $z(x, y)=\sqrt{1-x^{2}-y^{2}}$.
The illumination $\mathcal{I}$ is a dimensionless quantity normalized such that the
integral of $A\,\mathcal{I}$ over the units disc is equal to the flux measured by the
observer as a fraction of the flux of the illumination source.
To compute the total flux, we need to weigh each of the terms in the Green's basis
and integrate them over the visible portion of the body's disc to obtain
the reflected light solution vector $\mathbb{s}^\top$.
Evaluating these integrals is unfortunately intractable because of the piecewise
function in Equation~\ref{eq:illumination}. It is more tractable to weight the basis
terms by the function $I$ and modify the limits of integration such that the nightside
of the body is excluded. \citet{2021arXiv210306275L} shows that $I$ can be expressed
in the form of a linear operator $\mathbf{I}$ to weight a map vector  in the polynomial
basis $\tilde{\mathbf{p}}$ by the illumination profile. The total flux is then
\begin{equation}
    f=\mathbb{s}^\top\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right) \mathbf{A}_{\mathbf{2}} \mathbf{I}\left(b, \theta^{\prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime}\left(x_{\mathrm{o}}, y_{\mathrm{o}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}
\end{equation}
where
\begin{equation}
    \theta^{\prime}=\arctan 2\left(x_{\mathrm{o}}, y_{\mathrm{o}}\right)-\arctan 2\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right)
    \label{eq:starry_ref}
\end{equation}
is the angle of the terminator in frame $\mathcal{F}^\prime$.
In case the illuminated body is not occulted we have
\begin{equation}
    f_{0}=\mathbb{r}^{\top}(b) \mathbf{I}\left(b, \theta^{\prime \prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime \prime}\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}
    \label{eq:starry_ref_phase}
\end{equation}
where $\theta^{\prime\prime}=0$ is the angle of the terminator in frame
$\mathcal{F}^{\prime\prime}$ by construction. $\mathbf{R}^{\prime\prime}$ rotates the
body through an angle $\arctan 2\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right)$ so the
semi-major axis of the terminator is aligned with the $x^{\prime \prime}$ axis.
The solutions to integrals
$\mathbb{r}^{\top}(b)$ and $\mathbb{s}^{\top}\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right)$
are anything but trivial, they are derived in the appendices of
\citet{2021arXiv210306275L}. The key point is that the model is once again linear and
the matrices in Equations~\ref{eq:starry_ref} and \ref{eq:starry_ref_phase} need to be
computed only once for a specific occultation/illumination geometry.
\citet{2021arXiv210306275L} also derives a solution for the total flux in case the
light source is extended (the assumption that $r_s\gg 1$ is not satisfied) and
in case the body is not a perfect Lambertian reflector.

\subsection{The starry code}
The method described in the previous section is implemented in the
\textsf{Python} package
\textsf{starry}\footnote{\url{https://starry.readthedocs.io/en/latest/}}.
\textsf{starry} is many orders of magnitude faster and more accurate than
similar codes which rely on numerical integration. It enables computation of
rotational light curves, planetary transits, secondary eclipse light curves and
planet-planet occultations in both emitted and reflected light. Recent
extensions to \textsf{starry}, not mentioned in Section~\ref{sec:occultations},
extended the spherical harmonic formalism to the problem of Doppler imaging
\citep{2021arXiv211006271L} and add support for computing occultation light
curves of oblate bodies \citep{2022ApJ...925..185D}. In
Chapters~\ref{ch:mapping_io} and \ref{ch:mapping_exoplanets} we make extensive
use of \textsf{starry} to explore solutions to the \emph{inverse problem} --
how can we obtain an estimate of the map coefficients $\mathbf{y}$ which best
explain the observed light curve? \textsf{starry} abstracts the complexities of
computing the total flux and allows us to treat this problem as a linear
problem of the form
\begin{equation}
    f = \mathbf{X}\mathbf{y}
\end{equation}
where  $\mathbf{X}$ is computed by \textsf{starry} depending on the geometry of an
occultation or transit event. The inverse problem is highly degenerate
and making progress requires  imposing some prior structure on the solution
$\mathbf{y}$.

In Chapter~\ref{ch:mapping_io} we use \textsf{starry} to model occultation
light curves of a Solar System object -- Jupiter's moon Io, as it is occulted
by Jupiter. Thanks to the comparatively close distance to Io relative to
objects outside of the Solar System, these light curves are very high quality
we are able to fit very high order maps $l=20$ and test our models in a high
signal-to-noise regime. We also discuss how possible approaches to modeling
time-dependent maps in which case the spherical harmonic coefficients are
time-dependent: $\mathbf{y}=\mathbf{y}(t)$. In
Chapter~\ref{ch:mapping_exoplanets} we tackle a very similar problem but in a
very different regime of exoplanet eclipse mapping where the signal-to-noise is
orders of magnitude worse than in the case of Io. In both cases we mostly focus
on thermal light curves rather than light curves in reflected light.

\section{Bayesian statistics}
\label{sec:bayes}
Having covered the basic physics behind gravitational lensing and occultation mapping I
now turn to the subject of statistical inference in general and more focus more
specifically on Bayesian statistics in particular because it provides a straightforward
and elegant theoretical framework for solving inverse problems of the kind we mentioned
in Sections~\ref{sec:microlensing}  and \ref{sec:occultations}.

\subsection{Probability theory}
\subsection{The meaning of probability}
\subsection{Probability as frequency of events in repeated trials}
\subsection{Probability as degrees of belief}
\subsection{The likelihood function}
% mention the likelihood principle
\subsection{Priors}
\subsection{On machine learning and the difference between explanation and prediction}
% mention that even if we have a model focused on inference, it might be best if we judge 
% its quality through its predictive performance

\section{Inference}
\label{sec:inference}
\subsection{The curse of dimensionality}
% check that paper on physics and ML by that MIT guy, there's a good stuff there. 
% check also Ian Murray's lecture notes  on MLPR

\subsection{There's no free lunch}
% exploration vs. explotation 
\subsection{Optimization vs. sampling}
\subsection{Sampling and optimization methods}
\subsubsection{Maximum likelihood}
\subsubsection{The Laplace approximation}
\subsubsection{Variational inference}
\subsubsection{Importance sampling}
\subsubsection{Rejection sampling}
\subsubsection{Markov Chain Monte Carlo}
\subsubsection{Hamiltonian Monte Carlo}
\subsubsection{Other notable methods}
